\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size
\usepackage{./../usfassignment}
\settitle{Assignment 7}
\setauthor{Wanzhang Sheng}
\setcourse{CS673: Graduate Algorithms}

\begin{document}

\maketitle % Print the title

% -----------------------------------------------------------------------------
% PROBLEM 1
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  Let $p_1,p_2,\ldots,p_n$ be $n$ programs to be stored on a
  disk. Program $p_i$ requires $s_i$ kilobytes of storage, and the
  capacity of the disk is $D$, where $s_1+s_2+\cdots+s_n > D$.
\end{fancyquotes}

\begin{enumerate}
\item
  \begin{fancyquotes}
    (8 points) Give an efficient algorithm to determine which programs
    to store on the disk so that the total number of programs on the
    disk is maximized. If you give a greedy algorithm, be sure to
    prove that your algorithm is correct by proving both greedy choice
    and optimal program substructure. What is the running time of
    your algorithm, in terms of $n$?
  \end{fancyquotes}

  Greedy algorithm: pick the smallest program every time until no more
  can be put on the disk. Time cost in $O(n\lg{n})$.

  Prove:

  Greedy choice:
  There is at least one optimal solution which contain the greedy
  choice, putting the smallest one on the disk. Assume that there is
  no optimal solution contain the greedy choice. Pick a solution $S$,
  and replace any one $p_i$ by $p_1$(the smallest one) to get a new
  solution $S_2$. Since $p_1\leq p_i$ and $|S_1| = |S_2|$, $S_2$ is
  also a optimal solution. Contradiction. So there is at least one
  optimal solution contains the greedy choice.

  Optimal substructure:
  Assume there is no optimal solution contains the greedy choice but
  not optimal substructure.
  Let $S$ be the optimal solution, which contains $p_1$ but not
  contains optimal substructure for the rest $n-1$ programs
  pick. Generate $S_2$ by replacing the pick for the rest $n-1$
  programs pick by optimal substructure. Then $|S_1|=|S_2|$. Also
  substracture is optimal, $S_2$ is also less then $D$ space limits,
  which means it's a optimal solution. Contradiction.

  So greedy algorithm works.

  \begin{algorithm}[H]
    \SetKwProg{Fn}{Function}{}{end}
    \Fn{\textsc{GreedyPick}{(P,n,d)}}{
      P = qsort(P)\tcc*{By space size of the program in $O(n\ln{n})$}
      i = 0 \;
      sum = 0 \;
      set = \{\} \;
      \tcc{Assume P start with 0}
      \While{(i<n) and (sum + P[i].space $\leq$ d)}{
        sum += P[i].space \;
        set.add(P[i]) \;
        i +=1 \;
      }
      return set \;
    }
    \caption{Return the max number of program can be on the disk.}
  \end{algorithm}


\item
  \begin{fancyquotes}
    (8 points) Give an efficient algorithm to determine which
    programs to store on the disk so that as much of the disk as
    possible is utilized. If you give a greedy algorithm, be sure to
    prove that your algorithm is correct by proving both greedy
    choice and optimal program substructure. What is the running time
    of your algorithm, in terms of $n$?
  \end{fancyquotes}

  DP algorithm: Time cost in $O(nD)$.

  Table:
  $f(i,j), i\leq n, j\leq D$ indicated the max space usage with first
  $i$th programs and $j$ disk limits.

  Formula:
  \begin{equation*}
    f(i,j) =
    \begin{cases}
      0 & i\leq 0\ \text{or}\ j\leq 0\\
      \max(f(i-1,j), f(i-1,j-p_i)+p_i) & \text{otherwise}
    \end{cases}
  \end{equation*}

  Order:
  By $i$ from $0$ to $n$, then $j$ from $0$ to $D$.

  Code:
  \begin{algorithm}[H]
    \SetKwProg{Fn}{Function}{}{end}
    \Fn{\textsc{disk\_space}{(P, n, D)}}{
      F = array of n*D filled by 0 \;
      S = array of n*D filled by 0 \;
      \For{i = 1 $\rightarrow$ n}{
        \For{j = 1 $\rightarrow$ D}{
          \If{j >= P[i].space}{
            F[i, j] = F[i-1, j-P[i].space] + P[i].space \;
            S[i, j] = i \;
          }
          \If{F[i-1, j] > F[i, j]}{
            F[i, j] = F[i-1, j] \;
            S[i, j] = 0 \;
          }
        }
      }
      set = \{\} \;
      i = n \;
      j = D \;
      \While{i $>$ 0}{
        \If{S[i, j] > 0}{
          set.add(i) \;
          j -= P[i].space \;
        }
        i -= 1 \;
      }
      return set \;
    }
    \caption{Return the max space usage with given programs.}
  \end{algorithm}

\end{enumerate}
\pagebreak

% -----------------------------------------------------------------------------
% PROBLEM 2
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  (8 points) Problem 15--3 Edit Distance
\end{fancyquotes}

Table:
$f(i,j), 1\leq i\leq m, 1\leq j\leq n$ indicated the minimal cost to
transfer first $i$th chars of $x$ into $j$th chars of $y$.

Formula:
\begin{equation*}
  f(i,j) =
  \begin{cases}
    0 & i==0\ \text{and}\ j==0\\

    \min(f(i,j), f(i-1,j-1)+cost(copy)) & x[i]==y[j]\\
    \min(f(i,j), f(i-1,j-1)+cost(replace))\\
    \min(f(i,j), f(i-1,j)+cost(delete))\\
    \min(f(i,j), f(i,j-1)+cost(insert))\\
    \min(f(i,j), f(i-2,j-2)+cost(twiddle)) & x[i]==y[j-1],
    x[i-1]==y[j]\\
    \min(f(i,j), f(k,j)+cost(kill)) & 0\leq k\leq i-1, i==m, j==n\\
  \end{cases}
\end{equation*}

Order:
By $i$ from $0$ to $m$, then $j$ from $0$ to $n$.

Code:
\begin{algorithm}[H]
  \SetKwProg{Fn}{Function}{}{end}
  \Fn{\textsc{edit\_distance}{(X,m,Y,n)}}{
    F = array of m*n filled by infinity \;
    F[0, 0] = 0 \;
    \For{i = 0 $\rightarrow$ to m}{
      \For{j = 0 $\rightarrow$ to n}{
        \If{i $\geq$ 1, j $\geq$ 1, x[i]==y[j]}{
          F[i, j] = $\min$(F[i, j], F[i-1,j-1] + cost(copy))
        }
        \If{i $\geq$ 1, j $\geq$ 1}{
          F[i, j] = $\min$(F[i, j], F[i-1,j-1] + cost(replace))
        }
        \If{i $\geq$ 1}{
          F[i, j] = $\min$(F[i, j], F[i-1,j] + cost(delete))
        }
        \If{j $\geq$ 1}{
          F[i, j] = $\min$(F[i, j], F[i,j-1] + cost(insert))
        }
        \If{i $\geq$ 2, j $\geq$ 2, x[i]==y[j-1], x[i-1]==y[j]}{
          F[i, j] = $\min$(F[i, j], F[i-2,j-2] + cost(twiddle))
        }
      }
    }
    \For{k = 0 $\rightarrow$ m-1}{
      F[m, n] = $\min$(F[m, n], F[k,n] + cost(kill))
    }

    return F[m, n] \;
  }
  \caption{Return the minimal distance of edit from X to Y.}
\end{algorithm}

Time Cost: $O(mn)$

Space Cost: $O(mn)$

\pagebreak

% -----------------------------------------------------------------------------
% PROBLEM 3
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  (6 points) Exercise 16.2--5 Finding closed intervals

  Describe an efficient algorithm that, given a set $x_1, x_2, \ldots,
  x_n$ of points on the real line, determines the smallest set of
  unit-length closed intervals that contains all of the given
  points. Argue that your algorithm is correct.
\end{fancyquotes}

Greedy algorithm:
Sort the set in $O(n\lg{n})$. Find the most left point not covered
$x_i$, add interval $[x_i,x_i+1]$ into solution. Until all points are
covered, which the time cost is $O(n)$ since every point will be
accessed only once. The total time cost is $O(n\lg{n})$.

Greedy choice:
There is at least one optimal solution which contain the greedy
choice, $[x_{smallest},x_{smallest}+1]$. Assume there is no optimal
solution contains the greedy choice. Pick a optimal solution $S$,
replace the most left interval $p_1$ with $p_2 =
[x_{smallest},x_{smallest}+1]$ to get $S_2$. Since it's unit-length
interval, all points covered by $p_1$ will be covered by $p_2$. And
$|S|=|S_2|$, so $S_2$ is also a optimal solution.

Optimal substructure:
Assume there is no optimal solution contains the greedy choice but no
optimal substructure. Let $S$ be the optimal solution. After greedy
choice we covered first $j$ points, let the rest $n-j$ points are
covered by $k$ intervals. Consider a new solution $S_2$, the first $j$
points are samely covered by the greedy choice and the rest $n-j$
points are covered by optimal substructure. Then the new solution
$S_2$ is a feasible solution and since the rest points are covered by
optimal substructure solution, it will use at most $k$ intervals, so
the new solution $S_2$ is a optimal solution.

Greedy algorithm works.

Code:
\begin{algorithm}[H]
  \SetKwProg{Fn}{Function}{}{end}
  \Fn{\textsc{FCI}{(X, n)}}{
    set = \{\} \;
    A = qsort(X)\tcc*{time cost $O(n\lg{n})$}
    i = 1 \;
    \While{i $\leq$ n}{
      set.add([A[i]..A[i]+1]) \;
      t = A[i]+1 \;
      \While{A[i] $\leq$ t}{
        i += 1 \;
      }
    }
    return set \;
  }
  \caption{Return the }
\end{algorithm}

\pagebreak

\end{document}
