\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size
\usepackage{./../usfassignment}
\settitle{Assignment 5}
\setauthor{Wanzhang Sheng}
\setcourse{CS673: Graduate Algorithms}

\begin{document}

\maketitle % Print the title

% -----------------------------------------------------------------------------
% PROBLEM 1
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  (4 points) Does inserting and then immmediately deleting a unique
  element from a Binary Search Tree always leave the same tree? Does
  inserting and then immediatley deleting a unique element from a
  Red/Black tree always leave the same tree?
\end{fancyquotes}

Insert and then immediately delete a unique element in BST
\textbf{always} leave the same tree. Since the element is always be
inserted as a leaf, delete a leaf do not effect other nodes in the
BST.

Inserting and then immediately deleting a unique element in RBT
\textbf{may} change the tree. Since inserting and deleting both may
use rotating which is a action changed other nodes. Consider inserting
a node caused rotating but the element still a leaf, then after
deleting the tree will be different than before. But if there is only
one element in the tree and insert another one then delete it will let
it keep the same tree.

%\pagebreak

% -----------------------------------------------------------------------------
% PROBLEM 2
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  (8 points) Exercise 14.3-6 (MIN-GAP) Show how to extend a red-black
  tree to support the operation MIN-GAP, which gives the magnitude of
  the difference of the two closest numbers in the tree. For example,
  if $Q = \{1, 5, 9, 15, 18, 22\}$ then MIN-GAP($Q$) returns $18-15 = 3$
  since $15$ and $18$ are the two closest numbers in $Q$.
\end{fancyquotes}

Make $Q$ as a red-black tree $A$, then extend nodes of two more fields
for the minimal and maximal element of the subtree.

\begin{algorithm}[H]
  \SetKwProg{Fn}{Function}{}{end}
  \Fn{\textsc{min-gap}{(A, node)}}{
    return min(node.val - node.left.max, node.right.min - node.val,
    min-gap(A, node.left), min-gap(A, node.right))
  }
  \caption{Min-Gap of the given A in $O(n)$.}
\end{algorithm}

When inserting or deleting a node, we can update the fields of the
nodes in $O(\lg{n})$, by node.min = min(node.val, node.left.min) and
node.max = max(node.val, node.right.max). At the same time, let the
min and max fields of the null node be -inf and inf.

When we need to rotate the nodes, we can use the same method to
update the fields in $O(1)$ for each node, so that the rotating operation
will stay in $O(1)$.

%\pagebreak

% -----------------------------------------------------------------------------
% PROBLEM 3
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  (8 points) Show how to extend a red-black tree (that contains
  floating point numbers) so that you can calculate the average
  element (arithmetic mean) in constant time (while still taking
  $O(\lg{n})$ time for insert, find and delete).
\end{fancyquotes}

Extend two fields for the sum and the size of the subtree, then the
average number of the subtree is $\text{node.ave} = \text{node.sum} /
\text{node.size}$. And at the same time, no effect of find
operation. When inserting and deleting a element, update the nodes on
the path by $\text{node.sum} = \text{node.left.sum} +
\text{node.right.sum} + \text{node.val}$, $\text{node.size} =
\text{node.left.size} + \text{node.right.size} + 1$, which will let
the time cost stays in $O(\lg{n})$.

%\pagebreak

% -----------------------------------------------------------------------------
% PROBLEM 4
% -----------------------------------------------------------------------------
\section{}

\begin{fancyquotes}
  Problem 14-2 Josephus permutations The Josephus problem is defined
  as follows. Suppose that $n$ people are arranged in a circle and
  that we are given a positive number $m\leq n$. Beginning with a
  designated first person, we proceed around the circle, removing
  every $m$th person. After each person is removed, counting continues
  around the circle that remains. This process continues until all $n$
  people have been removed. The order in which the peope are removed
  from the circle defines the $(n,m)$-Josephus permitation of the
  integers $1,2,\ldots,n$. For example, the $(7,3)$-Josephus
  permutation is $<3,6,2,7,5,1,4>$.
\end{fancyquotes}

\begin{enumerate}
\item
  \begin{fancyquotes}
    (5 points) Suppose that $m$ is a constant. Describe an $O(n)$-time
    algorithm that, given an integer $n$, outputs the $(n,m)$-Josephus
    permutation.
  \end{fancyquotes}

  Just simulate the game process by link list, we can find the number
  for each time in $O(m)$ and delete it in $O(1)$. Since we need find
  $n$ times, the algorithm will run in $O(mn)$. If we treat $m$ as a
  constant, then the algorithm is $O(n)$.

\item
  \begin{fancyquotes}
    (5 points) Suppose that $m$ is not a constant. Describe an
    $O(n\lg{n})$-time algorithm that, given an integer $n$, outputs
    the $(n,m)$-Josephus permutation. Note that by not a constant
    here we mean that $m$ could be a function of $n$, like $n=2$,
    instead of a constant like $5$. $m$ still does not change over the
    course of the algorithm.
  \end{fancyquotes}

  We create a balanced tree for these $n$ numbers in $O(n\lg{n})$ and
  extend the nodes with a additional field indicated the size of the
  subtree in $O(n)$. So that we can determine the rank in the subtree
  of any node in $O(1)$ by $\text{node.rank} =
  \text{node.left.size}+1$. And find the $i$th element in $O(\lg{n})$
  by

  \begin{algorithm}[H]
    \SetKwProg{Fn}{Function}{}{end}
    \Fn{\textsc{find\_rank}{(node, i)}}{
      \eIf{node.rank = i}{
        return node
      }{
        \eIf{node.rank > i}{
          return find\_rank(node.left, i)
        }{
          return find\_rank(node.right, i - node.rank)
        }
      }
    }
    \caption{Find the $i$th element in a balanced tree.}
  \end{algorithm}

  Finding the next node by determine the rank $\text{next-rank} =
  ((\text{current-rank} + m -2) \bmod \text{current-size}) + 1$, and
  then find the node in $O(\lg{n})$. Using the equation $(a-1)\bmod n
  +1$ to change the result from $0\dots n-1$ to $1\dots n$. Since we
  need doing $n$ times, the final time cost will be $O(n\lg{n})$.

  Here is the math form:

  \begin{equation*}
    \text{rank}(i)=
    \begin{cases}
      1 & \text{if}\ i=0\ \text{or}\ i=n \\
      ((\text{rank}(i-1)+m-2) \bmod (n-(i-1))) +1 & \text{otherwise}
    \end{cases}
  \end{equation*}

\end{enumerate}


%\pagebreak

\end{document}
